Landscape composition and beneficial insects

### 2020-05-22 manuscript/manuscript_outline.Rmd
Added refs as sub-reference-database
- Tried using Water as a cover class, and it was highly "significant", but it gave some very strange results - not sure if model properly converged

---
title: "Landscape composition and beneficial insects"
author: "Sam Robinson, Paul Galpern, et al."
date: "Spring 2020"
output: bookdown::html_document2 
bibliography: "/home/samuel/Documents/bibtexRefs/references.bib"
csl: ecology-letters.csl
---

@@

* To characterize landscape composition surrounding the traps, we used classified landscape data from AAFC
  + Classifications are very accurate for fields (~90%), but less accurate for non-crop areas (~70%, [@AAFC_data])
  + Landscape studies often use the landscape composition within a circle of a fixed radius surrounding each point
  + However, the radius of the circle depends on the grain of the landscape that is relevant to the abundance of a given organism (@addicott1987)
  + Instead, we divided the landscape into annuli (rings) surrounding each trap location.
  + Each annulus was 30 m thick, and had inner radii ranging from 30 to 1470m, in increments of 30m.
  + Using functional regression (see below), we were able to incorporate landscape information at different distance, allowing us to assess the impact of both local and regional landscape composition

@@

* Scalar-on-function (SoF) regression is a special type of linear model ($\hat{y}=X\beta$), where the columns of the model matrix $X$ contain some continuous predictor of the scalar $y$ (e.g. biomass as a function of monthly average temperatures), and the values of the coefficients ($\beta$) are modeled as a smooth function $f(x)$ of the continuous predictors
  + In our case, each column represents the landscape composition within each annulus, with increasing radii away from the location of the trap, and the coefficients are a smooth function of distance ($\beta_i = f(distance_i)$), meaning that the coefficients represent the (additive) effect of a given type of landscape cover and  from the trap 
  + This avoids the fundamental problem of choosing an optimal radius of landscape composition, and allows the possibility that the size of the landscape ``grain" relevant to a given organism may change over the course of its life stages (@addicott1987, @lima1996)
  + This technique has been used by other authors (``onion-skin regression" in @galpern2020)
  + We fit our functional regression models using _mgcv_ version 1.8-31.

<!-- This part could be pared down. Probably not necessary to discuss several models if the full one is the only important one. -->
* We used 4 nested models to gauge the relative importance of landscape composition for arthropod activity density
  + Model 1 consisted of a temporal and spatial smoothers (``random effects"), necessary to account for the underlying spatiotemporal structure of the data 
  <!-- Smoothers can be thought of as random effects, where the coefficients follow a probability distribution defined by a fixed smoothing coefficient $\lambda$ - see @wood2017 for the Bayesian/Frequentist details of this -->
  + Model 2 added a local cover term, to account for the specific effect of local cover on activity density
  + Model 3 added a smoother term for the landscape rings, modeling the effect of non-crop cover at varying distances on activity density
  + Model 4, similar to model 3, modeled the effect of all underlying cover classes (crop and non-crop) at varying distances
  + To account for different lengths of trap exposure, log-days since trap placement was used as a fixed effect with its slope held at 1 (``offset" variable)
  + All models were fit using a negative binomial distribution, and model fits were compared using conditional Akaike's Information Criteria @wood2017

>>>>

---
title: "Landscape composition and beneficial insects"
author: "Sam Robinson, Paul Galpern, et al."
date: "Spring 2020"
output: bookdown::html_document2 
bibliography: refs.bib
csl: ecology-letters.csl
---

@@

* To characterize landscape composition surrounding the traps, we used classified landscape data from AAFC
  + Classifications are very accurate for fields (~90%), but less accurate for non-crop areas (~70%, [@AAFC_data])
  + Landscape studies often use the landscape composition within a circle of a fixed radius surrounding each point
  + However, the radius of the circle depends on the grain of the landscape that is relevant to the abundance of a given organism (@addicott1987)
  + Instead, we divided the landscape into annuli (rings) surrounding each trap location.
  + Each annulus was 30 m thick, and had inner radii ranging from 30 to 1470m, in increments of 30m.
  + Using functional regression [@ramsay2004], we were able to incorporate landscape information at different distance, allowing us to assess the impact of both local and regional landscape composition

@@

* We used functional regression to model the effect of landscape composition on activity density 
  +  Scalar-on-function (SoF) regression is a special type of functional linear regression model ($\hat{y}=X\beta$), where the columns of the model matrix $X$ contain some continuous predictor of the scalar $y$ (e.g. biomass in a plot (scalar) as a function of monthly average temperatures (vector)), and the values of the coefficients ($\beta$) are modeled as a smooth function $f(x)$ of the predictors
  + In our case, each column represents the proportion of a given landscape cover class within each annulus, with increasing radii away from the location of the trap, and the coefficients are a smooth function of distance ($\beta_i = f(distance_i)$), meaning that the coefficients represent the (additive) effect of a given type of landscape cover and  from the trap 
  + This avoids the fundamental problem of choosing an optimal radius of landscape composition, and allows the possibility that the size of the landscape ``grain" relevant to a given organism may change over the course of its life stages [@addicott1987, @lima1996]
  + This technique has been used by other authors [@yen2014, @galpern2020]
  + We fit our functional regression models using _mgcv_ version 1.8-31.

* Some of the landscape terms in our model were strongly concurved, so we removed them from the model, or combined them into a single term
  + Concurvity is an non-linear analogue of multicollinearity that can bias estimates of standard errors [@buja1989, @ramsay2003]
  + There is no agreed-upon threshold of unacceptable concurvity, but 0.5 is commonly used [@dominici2002, @ramsay2003] 
  + We found that some landscape terms were strongly concurved (Figure \@ref(fig:concurvity)), namely shrubland and forest (max: 0.71), canola and cereal (max: 0.74), and wetland and grassland (max: 0.63)
  +

```{r concurvity, echo=FALSE, fig.cap='\\label{fig:concurvity} Concurvity estimates for top land cover classes'}  
  include_graphics('../figures/coverCorPlots/concurvityEstimate.png')
```

<!-- This part could be pared down. Probably not necessary to discuss several models if the full one is the only important one. -->
* We used 4 nested models to gauge the relative importance of landscape composition for arthropod activity density
  + Model 1 consisted of a temporal and spatial smoothers (``random effects"), necessary to account for the underlying spatiotemporal structure of the data 
  + Smoothers can be thought of as random effects, where the coefficients follow a probability distribution defined by a fixed smoothing coefficient $\lambda$ [@wood2017] 
  + Model 2 added a local cover term, to account for the specific effect of local cover on activity density
  + Model 3 added a smoother term for the landscape rings, modeling the effect of non-crop cover at varying distances on activity density
  + Model 4, similar to model 3, modeled the effect of all underlying cover classes (crop and non-crop) at varying distances
  + To account for different lengths of trap exposure, log-days since trap placement was used as a fixed effect with its slope held at 1 (``offset" variable)
  + All models were fit using a negative binomial distribution, and model fits were compared using conditional Akaike's Information Criteria @wood2017
